### 2015-08-04 A4/writeup.tex
final commit

\section{Design}

>>>>

\section{Design}
This project is broken up into two separate pieces: a threaded portion and a multi-process portion. Both processes share certain functions, which are implemented almost identically: a sieve function which takes an integer input and marks as composite all multiples of that integer and a function which calculates whether an input integer is happy.\\

The sieve function is simple. It just modifies an integer array (memory mapped or normal). The happy function is also simple. It applies the happy number algorithm repeatedly and after each iteration adds the next number to a dynamic array. Once the next number is 1, it returns a 1. If the next number is ever found in the dynamic array, that means the cycle is looping and the number will never be happy, so it returns a 0.\\

The threaded program will take as input the number of threads to use and the maximum number for its range. This will allow for a script to dynamically time its behavior. The program will use mutexes and conditional variables. TLPI has a very good example in its second Threads chapter on how to use conditional variables to launch several threads, then stop them all. The example uses a conditional variable sort of like a semaphore. Every time a thread is about to die, it locks the mutex, decrements a thread counter, changes its state to terminated, unlocks the mutex and dies. Meanwhile the main thread wait()s on the worker threads. This wait functions works so that the main thread has the death mutex at all times except when it is waiting. Thusly threads can only die when the main thread is waiting for them to die. After the wait stops blocking, the main thread checks its children for termination. I will need to keep these threads working after they terminate, so I will have the program at this point refresh each thread and call it anew. I will repeat this loop until the Sieve of Eratosthenes has gone through all the numbers, up to the square root of the max prime. After this is done I will use the exact same method to figure out which primes are happy, except I will do no thread refreshing and instead just evenly distribute the work among all threads. After this I will somehow write all the happy primes to a file.\\

The multiprocess program will basically work the same way. It will do all the same behavior and use the same functions, slightly modified. Instead of making a bunch of new threads, it will fork(). The child process will do the child function (sieving or finding happiness). These functions will end with an exit call. To sieve, the program will start with two, have a child sieve that number, find the next prime, sieve on that number, etc etc until the square root of the max prime has been sieved. The sieve will then be complete. To make sure no more than PROCESSES processes are running, I will first make PROCESSES processes. I will have a data structure to count the aliveness / deadness of each child. I will then use the waitpid() function almost exactly like the conditional wait function of the threaded application to wait for processes to die, then renew them. In this way I will keep sieving until the primes are all found. I will then evenly distribute the work of calculating happiness among the child processes, write to file and be done. \\
